Traditional trade theory emphasizes trade creation (allowing cheaper
products from TA members to substitute for more expensive domestic
products) and trade diversion (substituting products from non-TA members
that were cheaper before the TA with products from TA members that are
cheaper now due to the TA reducing tariffs) (Schiff, Winters and Schiff,
2003) and argues that the impact of TAs depends on the comparative
advantage of member countries. In particular, it argues that TAs magnify
the impacts of a country's comparative advantage, relative to the world
and to other member countries signatories of a common TA. If member
countries of a TA have a comparative advantage on a factor endowment
relative to the world, but one country also has a comparative advantage
on the same factor endowment relative to the other member countries, the
country with the ``extreme'' advantage will be more vulnerable to trade
diversion effects, while countries with ``intermediate'' advantages will
gain from trade creation effects, predicting divergence of trade
outcomes, and winners and losers among member countries. (Venables,
2003). This emphasis on the trade creation and trade diversion effects
among member countries with significant differences in the comparative
advantage of their factor endowments relative to the world and to each
other, suggests that, when the country with the ``extreme'' comparative
advantage is a high-income country, relative to a lower-income country
with an ``intermediate'' comparative advantage, the lower-income country
should seek a TA with the other high-income country as it will gain
more. On the contrary, if both members are lower-income countries, the
country with the ``extreme'' comparative advantage, should not seek a TA
with the other low-income member country as it will be vulnerable.
(Sanguinetti, Siedschlag and Martincus, 2010). This logic can be easily
extended to the North-South and South-South types of TAs, as ``North''
countries will reasonably have an ``extreme'' comparative advantage in
skill-intensive goods relative to ``South'' countries, while ``South''
countries will reasonably have an ``extreme'' comparative advantage in
labour-intensive goods relative to ``North'' countries. Furthermore, it
is also argued in the literature that benefiting from economies of scale
through South-South economic integration is more difficult because
member countries do not have complementary production and trade
structures, nor high interpenetration of each other's markets on
intra-industry trade. (Schiff, Winters and Schiff, 2003). Also, South
countries can benefit from greater technological diffusion from
North-South TAs as the ``North'' countries have higher industrial
development as well as investment in research (Schiff and Wang, 2008).
Finally, as the trend in manufacturing has been in favour of vertical
specialization or value chain fragmentation (Krugman, 1995), North-South
TAs are preferable as developing countries strive to capture a greater
portion of the value added. Based on these arguments, developing
countries should therefore be better off entering into North-South
rather than South-South agreements.