The results of our extended model allowing for heterogenous effects of
TAs is shown in Appendix IV in Tables 20 through Table 24. Africa in
table 20 only has effects for South-South TAs and again has no
statistically significant effect for any TA. Americas in Table 21 has
five TAs with North-South estimates, one of which has statistically
significant and negative effects for NS TA + Lag and statistically
significant and positive effects for SS TA + Lag. Of the remaining four,
none have estimates for SS TA + Lag, three are statistically significant
and positive, and one is not statistically significant. It has eight TAs
with South-South estimates, seven of which have statistically
significant and positive effects, while one does not have statistically
significant effects. Americas does not have any coefficients for
North-North. Asia in Table 22 has two TAs with North-South estimates,
one of which is statistically significant and positive, while the other
is not statistically significant. It has nineteen TAs with South-South
estimates, seven of which have statistically significant and positive
effects, four have statistically significant and negative coefficients,
and eight does not have statistically significant effects. Asia does not
have any coefficients for North-North. Europe in Table 23 has eight TA
North-South estimates, five of which are statistically significant and
positive, and the others are not statistically significant. One of the
five agreements with statistically significant and positive coefficients
for NS TA + Lag also has a statistically significant and positive
coefficient for SS TA + Lag. None of the other agreements with a NS
coefficient have statistically significant coefficients for SS. It has
nineteen South-South estimates, thirteen are statistically significant
and positive, one is statistically significant and negative, and five
are not significant. Finally, the region has one agreement with a
North-North estimate, which also has a North-South and a South-South
estimate and they are all statistically significant and positive.
Intercontinental in Table 24 has thirty TA North-South estimates, of
which twelve are statistically significant and positive, fifteen are not
statistically significant, and three are statistically significant and
negative for NS TA + Lag. None of these TAs also have coefficients for
SS TA + Lag of which five are statistically significant and positive,
three are not statistically significant, and one is statistically
significant and negative. It has twenty-one estimates for South-South,
of which fourteen are statistically significant and positive, five are
not statistically significant, and two are statistically significant and
negative. It has three agreements with North-North estimates, two
statistically significant and positive, and one are not statistically
significant. Across the regions and TAs, 23 out of 47 (48.94\%) NS
coefficients have significant and positive effects, 20 out of 47
(42.55\%) NS coefficients have no significant effects, and 4 out of 47 (8.51\%) NS
coefficients have significand and negative effects; 49 out of 84 (58.33\%) SS coefficients
have significant and positive effects, 27 out of 84 (32.14\%) SS coefficients have no
significant effects, and 8 out of 84 (9.52\%) SS coefficients have significand and
negative effects; and, 3 out of 4 (75\%) NN coefficients have
significant and positive effects, 1 out of 4 (25\%) NN coefficients have no significant
effects, and none have significand and negative effects. A summary of
the findings can be found on Figure 6 for North-South trade, Figure 7 for
North-North trade and Figure 8 for South-South trade, with the
significance of the coefficients on the Y axis (all non-significant
coefficients assigned a value of -1 for ease of visualization, and
significant coefficients assigned a value of 1, 2 or 3 according to
their significance, with the highest significance being 3) magnitude of
the coefficients on the X axis, showing negative and positive
coefficients.
