The results of our model allowing for heterogenous effects of TAs is
shown in Tables 15 through Table 19 in Appendix III. Again, we can
observe significant heterogeneity across regions and TAs. Africa in
Table 15 has no statistically significant effect for any TA. Americas in
Table 16 has ten TAs with statistically significant and positive
coefficients, two with no statistically significant effect, and one TA
with a statistically significant and negative coefficient. Asia in Table
17 has eight TAs with statistically significant and positive
coefficients, nine with no statistically significant effect, and four
TAs with statistically significant and negative coefficients. Europe in
Table 18 has eighteen TAs with statistically significant and positive
coefficients, nine with no statistically significant effect, and one TA
with a statistically significant and negative coefficient. And finally,
Intercontinental in Table 19 has twenty-eight TAs with statistically
significant and positive coefficients, twenty with no statistically
significant effect, and six TAs with statistically significant and
negative coefficients. Across the regions, 64 out of 118 (54.24\%)
coefficients have significant and positive effects, 42 out of 118
(35.59\%) have no significant effects, and 12 out of 118 (10.17\%) have
significand and negative effects. A summary of the findings can be found
on Table 1, with the significance of the coefficients on the Y axis (all
non-significant coefficients assigned a value of -1 for ease of
visualization, and significant coefficients assigned a value of 1, 2 or
3 according to their significance, with the highest significance being
3) magnitude of the coefficients on the X axis, showing negative and
positive coefficients.
