Finally, we present the results of running our estimations substituting
trade flows as our dependent variable for the unit values of products
exported, specifically under the HS 2-digit codes 84 and 85 for
manufacturing products, in order to analyse if the effect of PTAs goes
beyond trade volumes. For ease of comparison, we ran each estimation
twice for each HS code: one with trade volume as the dependent variable,
and one with the unit value of the product exported as the dependent
variable.

Tables 13 and 14, and 15 and 16, show the results of our benchmark model
for each region for trade volumes and the unit value of the product
exported, and for HS 84 and 85, respectively. We continue to observe
heterogeneous results across regions. In table 13, for the trade volume
of HS 84, none of the PTA + Lag coefficients are statistically
significant with the exception of the Intercontinental region, for which
it is statistically significant and negative. In table 14, for the unit
value of the product exported of HS 84, the effects are not significant
for Africa and Asia, they are significant and negative for Americas, and
significant and positive for Europe and Intercontinental. Interestingly,
these results suggest that Intercontinental PTAs reduced the volume of
trade of HS 84 products but increased the value per unit. In table 15,
for the trade volume of HS 85, PTA + Lag coefficients are not
statistically significant for Americas, Asia and Intercontinental, while
Africa's results are significant and positive, and Europe's are
significant and negative. In table 16, for the unit value of the product
exported of HS 85, results are only slightly significant for
Intercontinental, with a negative coefficient. The rest of the regions
do not have significant results.

Tables 17 and 18, and 19 and 20, show the results of our extended
benchmark model with North-North, North-South and South-South PTAs, for
each region for trade volumes and the unit value of the product
exported, and for HS 84 and 85, respectively. In table 17, for the trade
volume of HS 84, we observe that for North-North trade, PTA + Lag
coefficient for Intercontinental has a significant and positive
coefficient, while Europe's is not significant. For North-South trade
PTA + Lag coefficients are not significant for Asia and Europe, while
they are significant and positive for Americas, and significant and
negative for Intercontinental. For South-South trade, PTA + Lag for
Africa, Asia and Europe do not have significant coefficients, while the
coefficients of Americas and Intercontinental are significant and
negative. In table 18, for the unit value of the product exported of HS
84, for North-North trade's PTA + Lag, Europe's coefficient is
significant and positive and the coefficient of Intercontinental is not
significant. For North-South trade, none of the PTA + Lag coefficients
are significant. For South-South trade, the PTA + Lag coefficients of
Africa, Americas and Asia are not significant, while Europe and
Intercontinental have significant and positive coefficients.
Interestingly, while trade volume for North-South and South-South for
Intercontinental PTAs decreased, the value per unit of South-South trade
increased. In table 19, for the trade volume of HS 85, we observe that
for North-North trade, PTA + Lag coefficient for Intercontinental has a
significant and positive coefficient, while Europe's is not significant.
For North-South trade PTA + Lag coefficients are not significant for
Americas, Asia and Intercontinental, while they are significant and
negative for Europe. For South-South trade, PTA + Lag for Americas,
Asia, Europe and Intercontinental do not have significant coefficients,
while the coefficient of Africa is significant and positive. In table
20, for the unit value of the product exported of HS 85, for North-North
trade's PTA + Lag, Europe and Intercontinental's coefficients are not
significant. For North-South trade, the PTA + Lag coefficients for
Americas and Europe are not significant, while they are significant and
negative for Asia and Intercontinental. For South-South trade, the PTA +
Lag coefficients of Africa, Americas and Intercontinental are not
significant, while Europe has significant and negative coefficients and
Asia has significant and positive coefficients. Interestingly, for
Asia's exports, the value per unit of product exported decreased with
North-South trade but increased with South-South trade.

Finally, for illustrative purposed, in tables 21 and 22, and 23 and 24,
we include the estimates of our model allowing for PTA specific effects,
extended with North-North, North-South and South-South PTAs, for Africa
and Americas, for trade volumes and the unit value of the product
exported, and for HS 84 and 85, respectively. In table 21, for the trade
volumes of HS 84 and 85 for Africa, which only has South-South PTAs, we
can see that PTA 670 had statistically significant and negative effects
on the trade volume of HS 84, and not significant for HS 85. PTA 787 did
not have a significant impact on trade volume of HS 84, while it has
significant and positive effects on HS 85. In table 22, for the unit
value of products HS 84 and 85 exported for the region of Africa, we can
see that PTA 670 did not have significant effects on the value per unit
of products in HS 84 and 85. PTA 787 did not have a significant impact
on the value per unit of HS 84, while it has significant and positive
effects on HS 85. This is a case where we can see a that a PTA has a
significant effect on the volume of trade and in the value per unit of a
category of manufacturing products of a South-South trade relationship.

In table 23, for the trade volumes of HS 84 and 85, and table 24 for the
unit value of products HS 84 and 85, all for the region of Americas,
which has North-South and South-South PTAs, we can observe heterogeneous
effects of different PTAs on the different types of bilateral trade
relationships. One interesting example is PTA 188, which has North-South
and South-South trade among its members. It has positive and significant
effects in the trade volumes of HS 84 and 85 for South-South trade,
while it has no significant effect in the trade volume of HS 84 and 85
for North-South trade. Furthermore, it has a significant and negative
effects on the value per unit of HS 84 for both North-South and
South-South trade, and it has no significant effect on the value per
unit of HS 85 for both North-South and South-South trade.
